### 2018-05-14 Docs/Thesis/docs/chapter1.tex
*1* ECLIPSE: OXYGEN. *2* MODIFICATION ON THE SIMITAR CODE. NOW RAPIDXML IS LOCATED IN DEPS DIR, AND THE CODE CHANGED A LITTE ON NETWORK TRACE CLASS.

Emerging technologies such as SDN and NFV can be considered great promises that would drastically change the development and operation of computer networks. However, its enabling technologies such as virtualization still pose challenges as for performance,  reliability, and security\cite{nfv-challenges}. Closed hardware solutions are easier to pass on Service Layer Agreement since they have a much more predictable behavior. Since it is expected that virtualization will affect performance negatively, these VNFs have to keep their degradation as small as possible. Therefore, guaranteeing the Service Layer Agreements becomes now a harder question. There is a demand for more reliable methods to ensure SLAs, over different types of loads.

@@

Furthermore realistic workload generators are an essential security research\cite{ditg-paper}, since generation of realistic workloads is important for evaluation of firewall middleboxes. It includes studies of intrusion, anomaly detection, and malicious workloads\cite{ditg-paper}. Over a virtualized enviroment a realistic workload is expected to have an even larger inpact.

Another critical point is the flow-oriented operation of SDN networks, in which each new flow arriving on an SDN switch demands an extra communication with the controller, that may cause a bottleneck. Since its operation relies on queries on flow tables, a stress load must have the same flow properties of an actual Internet Service Provider. Therefore, there is a demand for the study of the impact of a realistic traffic on this new sort of environment. How VNFs and virtualized middle-boxes and SDN testbeds will behave if stressed with a realistic traffic load in comparison to a constant rate traffic is a relevant subject.

@@

We also have a variety of APIs that enable the creation of traffic and custom packets, which include low-level APIs, such as Linux Socket API,  Libpcap, Libtins, DPDK, Pcapplusplus, libcrafter, impacket, scapy and many others. These APIs provide a finer controll and customization over each packet, and are used to implement traffic generators. For example, Ostinato and TCPreplay uses Libpcap, and MoonGen uses DPDK. In addition, many of the listed traffic generators provides their own API, such as Ostinato Python API, D-ITG C API, and MoonGen LUA API.

@@

There are few solutions on the open source community available that aiming to fill this gap, such as Harpoon\cite{harpoon-paper} and Swing\cite{swing-paper}.They capture traces to set internal parameters, providing an easier configuration. However they have their issues as well. Harpoon does not configure parameters at packet level\cite{harpoon-paper} and is not supported by newer Linux kernels, what may be a huge problem with setup and configuration. Swing\cite{swing-paper} aims to generate realistic traffic, but focous on background traffic; so throughputs it is not a goal off the application\cite{swing-paper} \cite{legotg-paper}. ecause its traffic generation engine is coupled to its modeling framework, you cannot opt to use a newer/faster packet generation library. The only way of replacing the traffic engine is changing and recompiling the original code which is a hard and  error prone activity\cite{legotg-paper}. Again, we fall in the same issue, a complex task that is usually not the goal of the project. In the table ~\ref{tab:related-work} we give a brief summary of what we have available on some open-source traffic-generators tools.

>>>>

Emerging technologies such as SDN and NFV can be considered great promises that are going to  drastically change the development and operation of computer networks if established in large scale. However, its enabling technologies such as virtualization still pose challenges as for performance,  reliability, and security\cite{nfv-challenges}. Closed hardware solutions are easier to pass on Service Layer Agreement since they have a much more predictable behavior. Since it is expected that virtualization will affect performance negatively, these VNFs have to keep their degradation as small as possible. Therefore, guaranteeing the Service Layer Agreements becomes now a harder question. There is a demand for more reliable methods to ensure SLAs, over different types of loads.

@@

Furthermore realistic traffic generators are an essential security research\cite{ditg-paper}, since generation of realistic workloads is important for evaluation of firewall middleboxes. It includes studies of intrusion, anomaly detection, and malicious workloads\cite{ditg-paper}. Over a virtualized enviroment a realistic workload is expected to have an even larger inpact.

Another critical point is the flow-oriented operation of SDN networks, in which each new flow arriving on an SDN switch demands an extra communication with the controller, that may cause a bottleneck. Since its operation relies on queries on flow tables, a stress load must have the same flow properties of an actual Internet Service Provider.

Therefore, there is a demand for the study of the impact of a realistic traffic on this new sort of environment. How VNFs and virtualized middle-boxes and SDN testbeds will behave if stressed with a realistic traffic load in comparison to a constant rate traffic is a relevant subject.

@@

We also have a variety of APIs that enable the creation of traffic and custom packets, which include low-level APIs, such as Linux Socket API,  Libpcap, Libtins, DPDK, Pcapplusplus, libcrafter, impacket, scapy and many others. These APIs provide a finer controll and customization over each packet, and are used to implement traffic generators. For example, Ostinato and TCPreplay uses Libpcap, and MoonGen uses DPDK. In addition, many of the listed traffic generators provides their own (high-level) APIs, such as Ostinato Python API, D-ITG C API, and MoonGen LUA API.

@@

There are few solutions on the open source community available that aiming to fill this gap, such as Harpoon\cite{harpoon-paper} and Swing\cite{swing-paper}.They capture traces to set internal parameters, providing an easier configuration. However they have their issues as well. Harpoon does not configure parameters at packet level\cite{harpoon-paper} and is not supported by newer Linux kernels, what may be a huge problem with setup and configuration. Swing\cite{swing-paper} only generates TCP and UDP traffic, and have limitations on throughputs rates\cite{swing-paper} \cite{legotg-paper}. Because its traffic generation engine is coupled to its modeling framework, you cannot opt to (easily) use a newer/faster packet generation library. The only way of replacing the traffic engine is changing and recompiling the original code which is a hard and  error prone activity\cite{legotg-paper}. Again, we fall in the same issue, a complex task that is usually not the goal of the project. In the table~\ref{tab:related-work} we give a brief summary of what we have available on some open-source traffic-generators tools.

### 2018-05-14 Docs/Thesis/docs/chapter2.tex
*1* ECLIPSE: OXYGEN. *2* MODIFICATION ON THE SIMITAR CODE. NOW RAPIDXML IS LOCATED IN DEPS DIR, AND THE CODE CHANGED A LITTE ON NETWORK TRACE CLASS.

The development of new network generators tightly related with to the current needs of network environment, applications sets, and purpose of use \cite{validate-trafficgen}. For example, if we want to evaluate the performance limits of equipment or new technology, you may want to execute stress tests, at specific constant rates. In this case, maximum throughput traffic generator is more interesting. On the other hand, to test how it performs on an actual network environment, you should search for a tool that produces a stochastically realist profile. Or, if you have a more specific usage case, you may search for a traffic generator that emulates an application scenario such as a client and a network server\cite{surge-paper}.

@@

There are many classifications of traffic generators\cite{sourcesonoff-paper}\cite{hybrid-traffic-gen}\cite{validate-trafficgen}\cite{do-you-trust}, and many have features that fall into more than one class. But classify them it is an efficient way to distinguish the difference between these tools. We will cite here two:

@@

The most common traffic generator classification in the literature is about its abstraction level\cite{do-you-trust}. There are four groups: Application-level, flow-level, packet-level and multi-level traffic generators. We present a diagram illustrating its concept in figure ~\ref{fig:layers-workload-tools}.

@@

\textbf{Flow-level traffic generators}: they are able to configure and reproduce features of flows\cite{do-you-trust}\cite{sourcesonoff-paper}, such as flow duration, start times distributions, and temporal (diurnal) traffic volumes\cite{do-you-trust}. Harpoon \cite{harpoon-paper} is able to extract these parameters from Cisco NetFlow data, collected from  routers.

\textbf{Packet-level traffic generators}: most traffic generators available fall in this class. They can craft and inject packets into the network, and control features like inter-departure times (IDT), packet size(PS), throughput and packets per second\cite{validate-trafficgen}. \cite{validate-trafficgen} classify them as replay engines, maximum throughput generators, and model-based generators. Traffic Replay engines, such TCPreplay\cite{web-tcpreplay}, can read packet capture files (\textit{pcap} files), and inject copies on a network interface. Model-based generators, such D-ITG \cite{ditg-paper}, TG\cite{web-tg}, may have PS and IDT configured by the user. They can configure IDT  either by constant values (defining the packet rate or bandwidth) or by stochastic models. Maximum throughput engines like Ostinato and Seagull are tools that the main goal is to perform end-to-end stress testing. Are included in this category generators that can craft packets from the link layer, like Brute, KUTE, and pktgen.

@@

Another desired characteristic is a high variability, in mathematical terms, an infinite variance. Process with such characteristic is said to be heavy-tailed \cite{sourcesonoff-paper}. In practical terms, that means a sudden discontinuous change can always occur. Heavy tail means that a stochastic distribution is not exponentially bounded\cite{sourcesonoff-paper}. This means that a value far from the mean does not have a negligible probability of occurrence. We can express self-similar and heavy-tailed processes using heavy-tailed stochastic distributions, such as Pareto and Weibull. As a reference for these stochastic distributions, a list in the table ~\ref{tab:distributions-equations}. In the last column, we indicate if the distribution is or not heavy-tailed.

@@

An important desirable feature that should be controlled is the packet size distribution since each flow may have its own packet distribution. As we can see in many works, the packet size of a trace may result in a huge impact in a trace throughput \cite{stochartic-selfsimilar}\cite{performance-trafficgen}; small packets cause a huge overhead in the packet throughput. In the table ~\ref{tab:packet-size-impact} there are a survey of results found by two different works \cite{comparative-trafficgen-tools} \cite{performance-trafficgen} about the impact of different packet sizes in the throughput rate.Therefore, it is an important factor we must taken into account in the evaluation of new proposals, and in the implementation of network workloads generators. On packet size distribution characterization, we can find many works in the literature. For example, \cite{packet-distribution-model} analyses many packet sizes distributions of many packet traces, in many environments. Some general results are that 90\% of UDP packets are smaller than 500 bytes, and most packets transmitted using TCP have 40 bytes (acknowledgment) and 1500 bytes (Maximum Transmission Unit, MTU) \cite{packet-distribution-model}. For UDP traffic, we have similar results, since mostly they are bimodal as well \cite{udp-flows-model}. Ostrowsky et al.\cite{udp-flows-model} found that, on UDP traces the modes of two regions are 120 and 1350 bytes, with a cut-off value of 750 bytes. They also find that roughly UDP packets constitute 20\% of the total number of packets on captures.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scaling Characteristics}

@@

On energy wavelet multiresolution energy plots, we can capture three different main behavior, according to the scale.  On \textbf{periodic time series}, the Energy values will be small. In fact, on perfectly periodic scales $j$, the values of the energy function $E_{j}$ will be zero. So periodicity will be sensed if the value of the energy function decrease. Perfect \textbf{white noise time series} will maintain the same value of the energy function. So an approximately constant values for the energy function $E_{j}$ indicates white noise behavior (which can be represented by a Poisson process\cite{poisson-white-noise}). On \textbf{self-similar time series}, the energy function plot $\log(E_{j})$ will grow approximately linearly with the scale $j$.

>>>>

The development of new network generators tightly related with to the current needs of network environment, applications sets, and purpose of use \cite{validate-trafficgen}. For example, if we want to evaluate the performance limits of equipment or new technology, you may want to execute stress tests, at specific constant rates. In this case, a maximum throughput traffic generator is more interesting\cite{validate-trafficgen}. On the other hand, to test how it performs on an actual network environment, you should search for a tool that produces a stochastically realist profile. Or, if you have a more specific usage case, you may search for a traffic generator that emulates an application scenario such as a client and a network server\cite{surge-paper}.

@@

There are many classifications of traffic generators\cite{sourcesonoff-paper}\cite{hybrid-traffic-gen}\cite{validate-trafficgen}\cite{do-you-trust}, and many have features that fall into more than one class. But classify them it is an efficient way to distinguish the difference between these tools. We will cite here two sorts of classification:

@@

The most common classification is about its abstraction level\cite{do-you-trust}, and it divides the traffic generators in four groups: Application-level, flow-level, packet-level and multi-level traffic generators. We present a diagram illustrating its concept in figure ~\ref{fig:layers-workload-tools}.

@@

\textbf{Flow-level traffic generators}: they can configure and reproduce flow characteristics\cite{do-you-trust}\cite{sourcesonoff-paper}, such as flow duration, start times distributions, and temporal (diurnal) traffic volumes\cite{do-you-trust}. Harpoon \cite{harpoon-paper} is able to extract these parameters from Cisco NetFlow data, collected from  routers.

\textbf{Packet-level traffic generators}: most common type of traffic generator. They can control packet-features like inter-departure times, packet size, throughput and packets per second\cite{validate-trafficgen}. \cite{validate-trafficgen} classify them as replay engines, maximum throughput generators, and model-based generators. Traffic Replay engines, such TCPreplay\cite{web-tcpreplay}, can read packet capture files (\textit{pcap} files), and inject copies on a network interface. Model-based generators, such D-ITG \cite{ditg-paper}, TG\cite{web-tg}, may have PS and IDT configured by the user. They can configure IDT  either by constant values (defining the packet rate or bandwidth) or by stochastic models. Maximum throughput engines like Ostinato and Seagull are tools that the main goal is to perform end-to-end stress testing. Are included in this category generators that can craft packets from the link layer, like Brute, KUTE, and pktgen.

@@

Another stochastic characteristic of the Ethernet traffic is high variability, or infinite variance. Processes with such characteristic are said to be heavy-tailed \cite{sourcesonoff-paper}. In practical terms, that means a sudden discontinuous change can always occur. Heavy tail means that a stochastic distribution is not exponentially bounded\cite{sourcesonoff-paper}. This means that a value far from the mean does not have a negligible probability of occurrence. We can express self-similar and heavy-tailed processes using heavy-tailed stochastic distributions, such as Pareto and Weibull. As a reference for these stochastic distributions, a list in the table ~\ref{tab:distributions-equations}. In the last column, we indicate if the distribution is or not heavy-tailed.

@@

Another relevant feature is the packet size distribution. As we can see in many works, the packet size of a trace may result in a huge impact in a trace throughput \cite{stochartic-selfsimilar}\cite{performance-trafficgen}, since small packets cause a larger overhead on packet processing. To exemplify, in the table~\ref{tab:packet-size-impact} we survey results from two different works \cite{comparative-trafficgen-tools} \cite{performance-trafficgen} about throughput impact of packet sizes. On packet size distribution characterization, we can find many works in the literature. For example, \cite{packet-distribution-model} analyses many packet sizes distributions of many packet traces, in many environments. Some general results are that \textbf{\textit{90\%}} of UDP packets are smaller than 500 bytes, and most packets transmitted using TCP have \textbf{\textit{40 bytes}} (acknowledgment) and \textbf{\textit{1500}} bytes (Maximum Transmission Unit, MTU) \cite{packet-distribution-model}. For UDP traffic, we have similar results, since mostly they are bimodal as well \cite{udp-flows-model}. Ostrowsky et al.\cite{udp-flows-model} found that, on UDP traces the modes of two regions are \textbf{\textit{120}} and \textbf{\textit{1350}} bytes, with a \textbf{\textit{cut-off}} value of \textbf{\textit{750 bytes}}. They also find that roughly UDP packets constitute \textbf{\textit{20\%}} of the total number of packets on captures.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fractal and Scaling Characteristics}

@@

On energy wavelet multiresolution energy plots, we can capture three different main behavior, according to the scale.  On \textbf{periodic time series}, the Energy values will be small. In fact, on perfectly periodic scales $j$, the values of the energy function $E_{j}$ will be zero. So periodicity will be sensed if the value of the energy function decrease. Perfect \textbf{white noise time series} will maintain the same value of the energy function. So an approximately constant values for the energy function $E_{j}$ indicates white noise behavior (which can be represented by a Poisson process\cite{poisson-white-noise}). On \textbf{self-similar time series}, the energy function plot $\log(E_{j})$ will grow approximately linearly with the scale $j$.  This explains why it has become a standard for realistic traffic generation analysis. In a single plot, we can easily identifiy periodicities, and self-similar and Poisson process characteristics, just seeing if it decays, grow, or remain constant.

### 2018-05-14 Docs/Thesis/docs/chapter3.tex
*1* ECLIPSE: OXYGEN. *2* MODIFICATION ON THE SIMITAR CODE. NOW RAPIDXML IS LOCATED IN DEPS DIR, AND THE CODE CHANGED A LITTE ON NETWORK TRACE CLASS.

In this chapter, we will present our tool which aims to fill the gaps addressed in the chapter ~\ref{ch:introduction}. We call it \textit{SIMITAR}, an acronym for \textit{SnIffing, ModellIng and TrAffic geneRation}.  This acronym summarizes its operation. SIMITAR is a traffic generator able to learn features of real traffic automatically, and reproduce synthetic traffic similar to the original. It records a model for the traffic in an XML file we call \textit{compact trace descriptor} (CTD file). As input-data SIMITAR can use \textit{pcap} files or real-time captures.

@@

On the figure ~\ref{fig:layer-diagram} we abstract our concepts we stated in a layer diagram model. Or tool must work as an interface for the user automatizing traffic modeling, configuration, and emulation and offer services of intelligence and traffic generation programmability. In the figure, we also include packet-acceleration\footnote{Packet acceleration is a concept introduced by DPDK\cite{web-dpdk}, which means kernel by-pass. Packet acceleration optimizes the packet processing, and therefore traffic generation, enabling higher throughput rates.}, however, in the current version, this feature is not implemented, but how to do it is discussed in the section of future works in the chapter ~\ref{ch:conclusion}. To avoid confusions, we will refer to the underlying workload tool as packet generator engine/tool, and for SIMITAR as a traffic generator. Even if this packet generator may be a traffic generator itself (such as Iperf), on SIMITAR, its operates just on the packet-level, as we explain in the chapter ~\ref{ch:literature-review}, and our tool is a multi-layer traffic generator tool.

Our primary goal is to offer simple configuration, with \textbf{realism}, at a reasonable speed. The intermediate layer of the figure ~\ref{fig:layer-diagram} summarize, the purpose of the project in an illustrative way. Instead of storing a huge file \textit{pcap} with a size of many \textit{gigas}, we create a light-weight set of models and parameters that describe this same trace. We are also introducing the concept of \textbf{programmability}. Since the \textit{compact trace descriptor},  is an XML file, the user may create his custom traffic in a platform agnostic way creating his own CTD file. So the user will be able to create multi-flow Ethernet traffic not having to build many scripts for each tool or reading its documentation. A single description of the traffic will fit any supported packet generator. Using a component methodology, we decouple the packet generation, from the data collection and parameterization process. We developt it using the design pattern factory to make the extension easy for any packet generator engine.

We abstract its whole operation cycle in the figure ~\ref{fig:cycle-of-operation}. Our tool, from live captures or \textit{pcap} file collects raw data from Ethernet traffic. It then breaks these data into different flows and uses this data to generate a set of parameters for our traffic model, using a set of algorithms. Finally, SIMITAR provides these parameters to a packet generator engine and controls its packets injection.

@@

To meet the requirements presented in the chapter ~\ref{ch:introduction}, we will define our solution, and how it works. SIMITAR architecture is presented in the figure~\ref{fig:architecture}. It is composed of four components: a \textit{Sniffer}, a \textit{SQLite database}, a \textit{TraceAnalyzer}, a \textit{FlowGenerator};  and a \textit{Network Traffic Generator}, as subsystem. We describe each part below.

@@

This component collects network traffic data and classifies it into flows, storing stores in an SQLite database. It defines each flow by the same criteria used by SDN switches\cite{sdn-survey},  through header fields matches. It uses:

@@

It has a data structure we developed called \textit{OrderedSet}. A set is a list of elements with no repetition but does not keep track of the insertion order. Our OrderedSet does. Also, it makes use of a  64 bits hash function of the family FNV\footnote{The collision probability of a good 64 bits hash function in a table with 10000 items is about of $2.71e-12$.}. The listed header fields are inputs for a hash function. The hash value is then added to the ordered set which returns its order (index on the \textit{OrderedSet}). The index value is chosen as a packet flowID.

@@

\begin{figure*}[ht!]
    \centering
    \includegraphics[height=2.0in]{figures/ch3/modified-harpoon-model}
    \caption{The schema of the modified version of the Harpoon algorithm we adopt on SCIMITAR.}
    \label{fig:modified-harpoon-model}
\end{figure*}

To represent inter-packet times, we adopt a modified version of the Harpoon's traffic model. A deep explanation of the original model can be found at \cite{harpoon-paper} and \cite{harpoon-validation}. Harpoon uses a definition of each level, based on the measurement of SYN and ACK TCP flags. It uses TCP flags (SYN) to classify packets in different levels, named their file, session, and user level. We choose to estimate these values, based on inter-packet times only. The distinction is made based on the time delay between packets. For our purposes, the advantage of this strategy is that the evaluation of the packet-train is not attached to a specific protocol. We illustrate this behavior in the figure~\ref{fig:modified-harpoon-model}.

@@

In the \textbf{flow layer} the \textit{TraceAnalyzer} loads the flow arrival times from the database and calculates the inter-packet times within the flow context. At the \textbf{session layer}, we use a deterministic approach for evaluating file transference time and times between files: ON/OFF times sequence of packet trains. We choose a deterministic model because in this way we can express diurnal behavior\cite{harpoon-paper}.  We develop an algorithm called \textit{calcOnOff} responsible for estimating these times. It also determines the number of packets and bytes transferred by each file. Since the ON times will serve as input for actual traffic generators, we defined a minimum acceptable time for on periods equals to 100 ms. ON times can be arbitrary smalls, and they could be incompatible with acceptable ON periods for traffic generators. Also in the case of just one packet, the ON time would be zero. So setting a minimum acceptable time to solve these issues. The OFF times, on the other hand, are defined by the constant \texttt{session\_cut\_time} \footnote{In the code it is called \texttt{DataProcessor::m\_session\_cut\_time} }. If the time between two packets of the same flow is larger than \texttt{session\_cut\_time}, we consider them belonging to a different file, so this time is a session OFF time. In this case, we use the same value of the constant \textit{Request and Response timeout} of Swing\cite{swing-paper} for the \texttt{session\_cut\_time}: 30 seconds. The control of ON/OFF periods in the traffic generation is made by the \textit{Flow Generator} component \footnote{This control is made by the class \texttt{NetworkFlow}}.

In the \textbf{file layer}, we model the inter-packet times at the file level. To estimate inter-packet times within files, we select all inter-packet times smaller than \texttt{session\_cut\_time}\footnote{In the code it is called\texttt{DataProcessor::m\_session\_cut\_time} }. All files within the same flow are considered to follow the same model. We delegate the control of the inter-packet times to the underlying packet generator engine. We ordered them, from the best to the worst. Currently, we are using eight different stochastic functions parameterizations. They are Weibull(linear regression), Normal(mean/standard deviation calculation), Exponential(mean and linear regression estimation), Pareto(linear regression and maximum likelihood), Cauchy(linear regression) and Constant(mean calculation). From those, Weibull, Pareto, and Cauchy are heavy-tailed functions, and therefore self-similar processes. But if the flow has less than 30 packets, just the constant model is evaluated. It is because numerical methods gave poor results if the data sample used is small. We sort these models according to the Akaike Information Criterion (AIC) as default\cite{sourcesonoff-paper}\cite{bic-aic-comparision}. We will enter into deeper details on this methodology on the chapter ~\ref{ch:modeling-evaluation}. We present the methodology of selection in the figure~\ref{fig:model-parameterization}. All these constants and modes of operation are modifiable via command-line options.

>>>>

In this chapter, we will present our tool which aims to fill the gaps addressed in the chapter~\ref{ch:introduction}. We call it \textit{SIMITAR}, an acronym for \textit{SnIffing, ModellIng and TrAffic geneRation}\footnote{And Scimitar is curved sword originated on the Middle East.}.  This acronym summarizes its operation. SIMITAR is a traffic generator able to learn features of real traffic automatically, and reproduce synthetic traffic similar to the original. It records a model for the traffic in an XML file we call \textit{compact trace descriptor} (CTD file). As input-data SIMITAR can use \textit{pcap} files or real-time captures.

@@

On the figure ~\ref{fig:layer-diagram} we abstract our concepts we stated in a layer diagram model. Or tool work as an interface for the user automatizing traffic modeling, configuration, and emulation and offer services of intelligence and traffic generation programmability. In the figure, we also include packet-acceleration\footnote{Packet acceleration is a concept introduced by DPDK\cite{web-dpdk}, which means kernel by-pass. Packet acceleration optimizes the packet processing, and therefore traffic generation, enabling higher throughput rates.}, however, in the current version, this feature is not implemented, but how to do it is discussed in the section of future works in the chapter ~\ref{ch:conclusion}. To avoid confusions, we will refer to the underlying workload tool as packet generator engine/tool, and for SIMITAR as a traffic generator. Even if this packet generator may be a traffic generator itself (such as Iperf), on SIMITAR, its operates just on the packet-level, as we explain in the chapter ~\ref{ch:literature-review}, and our tool is a multi-layer traffic generator tool.

Our primary goal is to offer all services mentioned in the table~\ref{tab:related-work}, realismo on the synthetic traffic, auto-modelling and traffic configuration, simple usage,  configuration, and plain possibility of traffic customization for the user. The intermediate layer of the figure ~\ref{fig:layer-diagram} summarize, the purpose of the project in an illustrative way. Instead of storing a huge file \textit{pcap} with a size of many \textit{gigas}, we create a light-weight set of models and parameters that describe this same trace. We are also introducing the concept of \textbf{programmability}. Since the \textit{compact trace descriptor}(CDF), is an XML file, the user may create his custom traffic in a platform agnostic way creating his own CTD file. So the user will be able to create multi-flow Ethernet traffic not having to build many scripts for each tool or reading its documentation. A single description of the traffic will fit any supported packet generator. Using a component methodology, we decouple the packet generation, from the data collection and parameterization process. We developt it using the design pattern factory to make the extension easy for any packet generator engine.

We abstract its whole operation cycle in the figure ~\ref{fig:cycle-of-operation}. Our tool collect packet data  from live captures or \textit{pcap} files. It then breaks the traffic into flows and uses the data to generate  parameters for our traffic model. Finally, SIMITAR provides these parameters to a packet generator engine and controls the packet injection.

@@

To meet the requirements presented in the chapter ~\ref{ch:introduction}, we will define our solution, and how it works. SIMITAR architecture is presented in the figure~\ref{fig:architecture}. It is composed of four components: a \textit{Sniffer}, a \textit{SQLite database}, a \textit{TraceAnalyzer}, a \textit{FlowGenerator},  and a \textit{Network Traffic Generator}, as subsystem. We describe each part below.

@@

This component collects network traffic data and classifies it into flows, storing stores in an SQLite database. It uses header field matches to classify the flows, such as in SDN switches\cite{sdn-survey}. The used fields are:

@@

It has a data structure we developed called \textit{OrderedSet}. A set is a list of elements with no repetition but does not keep track of the insertion order. Our OrderedSet does. Also, it uses a  64 bits hash function from the FNV\footnote{The collision probability of a good 64 bits hash function in a table with 10000 items is about of $2.71e-12$.} family. The listed header fields are inputs for a hash function. The hash value is then added to the ordered set which returns its order (index on the \textit{OrderedSet}). The index value is chosen as a packet flowID.

@@

\begin{figure*}[ht!]
    \centering
    \includegraphics[height=2.0in]{figures/ch3/modified-harpoon-model}
    \caption{The schema of the modified version of the Harpoon algorithm we adopt on SIMITAR.}
    \label{fig:modified-harpoon-model}
\end{figure*}

To represent inter-packet times, we adopt a modified version of the Harpoon's traffic model. A deep explanation of the original model can be found at \cite{harpoon-paper} and \cite{harpoon-validation}. Harpoon uses a definition of each level, based on the measurement of SYN and ACK TCP flags. It uses TCP flags (SYN) to classify packets in different levels, named their file, session, and user level. We choose to estimate these values, based on inter-packet times only. The distinction is made based on the time delay between packets. For our purposes, the advantage of our strategy is that the evaluation of the packet-train is not attached to a specific header field, so we can define packet trains for any sort of flow. We illustrate this behavior in the figure~\ref{fig:modified-harpoon-model}.

@@

In the \textbf{flow layer} the \textit{TraceAnalyzer} loads the flow arrival times from the database and calculates the inter-packet times within the flow context. At the \textbf{session layer}, we use a deterministic approach for evaluating file transference time and times between files: ON/OFF times sequence of packet trains. We choose a deterministic model because in this way we can express diurnal behavior\cite{harpoon-paper}.  We develop an algorithm called \textit{calcOnOff} that estimate these times. It also determines the number of packets and bytes transferred by each file. Since the ON times will serve as input for actual traffic generators, we defined a minimum acceptable time for on periods equals to 100 ms. ON times can be arbitrary smalls, and they could be incompatible with acceptable ON periods for traffic generators. Also in the case of just one packet, the ON time would be zero. So setting a minimum acceptable time to solve these issues. The OFF times, on the other hand, are defined by the constant \texttt{session\_cut\_time} \footnote{In the code it is called \texttt{DataProcessor::m\_session\_cut\_time} }. If the time between two packets of the same flow is larger than \texttt{session\_cut\_time}, we consider them belonging to a different file, so this time is a session OFF time. In this case, we use the same value of the constant \textit{Request and Response timeout} of Swing\cite{swing-paper} for the \texttt{session\_cut\_time}: 30 seconds. The control of ON/OFF periods in the traffic generation is made by the \textit{Flow Generator} component \footnote{This control is made by the class \texttt{NetworkFlow}}.

In the \textbf{file layer}, we model the inter-packet times at the file level. We select all times smaller than \texttt{session\_cut\_time}\footnote{In the code it is called\texttt{DataProcessor::m\_session\_cut\_time} }, and all files within the same flow are considered to follow the same model. We delegate the control of the inter-packet times to the underlying packet generator engine. We ordered them, from the best to the worst. Currently, we are using eight different stochastic functions parameterizations. We display each of them on the table~\ref{tab:parameterizations-sumary} .

\begin{table}[ht!]
	\centering
	\caption{Functions and parameterizations used by SIMITAR}
		\begin{tabular}{c c c c}
			\toprule
			\textbf{Function} & Linear Regression & Maximum Likelihood & Empirical\footnotemark \\
			\midrule
			Weibull         & $\checkmark$    &                  &                  \\
			Normal             &                  &                  &  $\checkmark$    \\
			Exponential      & $\checkmark$    &                  &  $\checkmark$    \\
			Pareto          & $\checkmark$    & $\checkmark$    &                  \\
			Cauchy          & $\checkmark$    &                  &                  \\
			Constant          &                 &                  & $\checkmark$    \\
			\bottomrule
		\end{tabular}
		\label{tab:parameterizations-sumary}
\end{table}
\footnotetext{Empirical estimation, by calculation of the avarge, and standard deviation}

From the functions presented in the first column~\ref{tab:parameterizations-sumary}, Weibull, Pareto, and Cauchy are heavy-tailed (and self-similar processes). But if the flow has less than 30 packets, just the constant model is evaluated. It is because numerical methods gave poor results if the data sample used is small. We sort these models according to the Akaike Information Criterion (AIC) as default\cite{sourcesonoff-paper}\cite{bic-aic-comparision}. We will enter into deeper details on this methodology on the chapter ~\ref{ch:modeling-evaluation}. We present the methodology of selection in the figure~\ref{fig:model-parameterization}. All these constants and modes of operation are modifiable via command-line options.
